---
title: "STAT 425 Project 02"
author: "Daniel Girvitz & Ethan Scott"
date: "24/02/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("C:/Users/ethan/OneDrive/Documents/Stat 425/Project #2")
#setwd("C:/Users/Daniel/Documents/School/Uni/Fourth Year/STAT 425/Project 02")
```

### Preliminary

This section serves to demonstrate that regular R functions handle BIBDs correctly. We begin with the analysis of data in Table 4.22.

```{r}

time <- c(73,74,   71,
             75,67,72,
          73,75,68,  
          75,   72,75)
treatment = c(1,1,1,
              2,2,2,
              3,3,3,
              4,4,4) #catalyst
block = c(1,2,4,
          2,3,4,
          1,2,3,
          1,3,4) #batch

df <- data.frame(time, treatment, block)

library(crossdes)
(tab <- xtabs(~ treatment + block, data = df))
(d <- t(apply(tab, 1, function(x) (1:4)[x != 0])))
isGYD(d)

```

We have verified this data is a BIBD.

```{r}
one.way <- aov(time~factor(treatment)+factor(block), data=df)
summary(one.way)
```

As the ANOVA table does not match Table 4.24, the 'aov' function fails to analyze a BIBD properly. Thus, we must do this manually...

```{r}
test = aov(time~factor(block)+factor(treatment), data=df)
summary(test)
```

This one matches the Table

```{r, echo=FALSE}

#a = 4 # treatments
#b = 4 # blocks
#k = 3 # number of treatments in each block
#r = 3 # number of times each treatment occurs (in a block)

#N = 12
#a*r
#b*k

#The equation $N=ar=bk$ is satisfied.
```

### The data... along with some data cleaning

```{r}
# load data
eData.raw <- read.csv(file = 'appletaste.csv')
aftertaste <- eData.raw$aftertaste
panelist <- eData.raw$panelist
product <- as.factor(eData.raw$product)
eData <- data.frame(aftertaste,
                    panelist,
                    product)
```

### Checking for a balanced design

```{r}
# check balance for panelists
replications(aftertaste ~ panelist, data=eData)

# check balance for product
replications(aftertaste ~ product, data=eData)

# let's try a table
table(eData$panelist,eData$product)
```

Each of the twenty panelists samples three of the four products. Each product is sampled 15 times. Thus, the design appears to be balanced.

### Boxplots

```{r}
boxplot(aftertaste~product,data=eData, 
        main="Aftertaste data",
        xlab="Product", 
        ylab="Aftertaste")
stripchart(aftertaste ~ product, data=eData,
           vertical = TRUE, 
           method = "jitter",
           pch = 20, 
           add = TRUE, 
           col = "rosybrown1")
means <- tapply(eData$aftertaste, eData$product, mean)
points(means, pch=19, col="red")
```

Product 493 appears to be the sweetest. The rest are relatively more bitter; as regards their differences, that is TBD.

### ANOVA

```{r}
appletaste.aov <- aov(aftertaste ~ product + panelist, data=eData)
summary(appletaste.aov)
```

Obviously, the null hypothesis of equality of treatment means, holding a level of alpha=0.05, must be REJECTED.

##Below is the correct ANOVA table, the conclusion that can be drawn is the same

```{r}
appletaste.aov=aov(aftertaste~panelist+product, data=eData)
summary(appletaste.aov)
```

```{r}
par(mfrow=c(1,2))
plot(appletaste.aov,1)
plot(appletaste.aov,2)
```

\textbf{The Residuals vs Fitted plot:} Residuals are randomly distributed with what appears to be constant width, thus, we can conclude the homogenity of variance (homoscedasticity) requirement is satisfied. Although there is a curve to the red line, we choose to ignore it, as it is quite minor. A Box-Cox transformation may correct this slight deviation. \newline

\textbf{Normal Q-Q:} Theoretical quantities fit the diagonal very well, which intersects the origin. Thus, the normality requirement is satisfied. But, just to be sure, let's check the residuals. \newline

```{r}
par(mfrow=c(1,2))
eij=residuals(appletaste.aov)
hist(eij,main="Histogram of residuals")
plot(density(eij),main="Density plot of residuals",ylab="Density",xlab="Residuals")
```

Beauty. Perfect.

```{r}
shapiro.test(eData$aftertaste)
# bartlett.test(aftertaste ~ product + panelist, data=eData) Fails?
```

```{r}
library(DescTools)
LeveneTest(aftertaste~panelist, data=eData)
LeveneTest(aftertaste~product, data=eData)
```

Idk if this is the correct way but both p-values are above 0.05 so homogeneity of Variance holds

#Comparing Treatment means using Fisher LSD

```{r}
PostHocTest(appletaste.aov, method="lsd")
```

#Comparing means using Tukey HSD

```{r}
PostHocTest(appletaste.aov, method="hsd")
```
